Cultural Transmission and Technological Transitions during the Late Paleolithic in Korea

### 2022-08-22 analysis/paper/paper-doc.qmd
edit pca plots

@@

![Boxplot for each attribute measured on the stemmed points. The black bold horizontal lines in the middle of boxes represent the median value.](images/003-cv-sharma.png){#fig-three width="397"}

>>>>

1.  Distribution of attributes

@@

![Boxplot for each attribute measured on the stemmed points. The black bold horizontal lines in the middle of boxes represent the median value.](images/002-attributes-variation-01.png){#fig-three width="324"}

2.  Coefficient of variation

We used CV to measure the variations between stemmed points. If the variation is high, the transmission might happen through guided variation while the low variation represents indirect bias. Our results show that CV values for all attributes are distributed from 23.8 to 36.3 (@fig-four). Compared to the CVs of Great Basin projectile points [@eerkens2001techniques], the average CV is higher. Unlike the distribution of actual attributes (@fig-three), the CV of body length and maximum length have the lowest values. The lower CV of maximum length indicates highly standardized with narrowest confidence intervals. Tang and stem related attributes (e.g. SL, SW, TL, TW) are less standardized with higher CV values and wider confidence intervals than other attributes on the artifacts. Among all the attributes, only maximum length has the CV values lower than 25, which could be the result of indirect bias.

![The points and blue colored numbers represent CV values for each attribute measured on the stemmed points. 95% confidence intervals for the CVs were computed using Sharma & Krishna's method. The vertical lines indicate the range of confidence intervals.](images/003-cv-sharma-01.png){#fig-four width="339"}

3.  Principal Component Analysis

PCA serves to examine correlation of attributes for determining the dominant mode of cultural transmission.Figure 5 shows positive relationship between attributes. Especially, there are strong relationships between body length and maximum length, tang width and mid width, and tang length and stem length. Body length and maximum length contribute the most to the shape of stemmed points and they are most correlated to each other. Other width and tang related attributes are close to each other but a bit far from body length and maximum length. Stem width contributes the least to the overall shape. We can assume that
